---
title: "Mapping FARS Data"
author: "Tom Mullie"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping FARS Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(magrittr)
library(dplyr)
library(tidyr)
library(maps)
library(graphics)
source("../R/fars_functions.R")
```
This package is designed to analyse and plot data from the **Fatality Analysis Reporting System**, or **FARS**. Information about the FARS system is available [at this address](https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)). 

## FARS Data
This package is designed to parse FARS data stored by year in comma separated values format. The data should be saved in the working directory.

## Dependencies
This package requires several packages to be available in order to funcion:
* magrittr
* dplyr
* tidyr
* maps
* graphics

## Using the Package
Two functions in the package will be of interest to users: "fars_summarize_years(years)"" and "fars_map_state(state.num, year)". Other functions, "fars_read"", "make_filename"" and "fars_read_years" are made available for the purposes of error checking, but will be of little interest in terms of data analysis. 

## Analysing FARS Incident Counts
The "fars_summarize_years(years)"" function produces a summary table of the number of FARS incidents in each month of a given year or group of years. For example, if data were available for 2013, 2014 and 2015, a summary table of incidents by month in each of those years could be generated by:
```{r echo=TRUE}
fars_summarize_years(c(2013, 2014, 2015))
```

## Mapping FARS Incidents in a State
The "fars_map_state(state.num, year)"" produces a map of all FARS incidents in a state in a given year. For example, to plot all of the FARS incidents in California in 2014, use the following code:
```{r eval=FALSE}
fars_map_state(6,2014)
```
Which should produce the plot below:

```{r echo=FALSE}
fars_map_state(6,2014)
```

A vector map of California, with dots indicating the locatoin of each FARS incident. 
